---
title: "Quiz 1: Part 2"
format: html
editor: visual
embed-resources: true
css: |
      <style>
        body { font-family: Arial, sans-serif; }
      </style>
js: |
      <script>
        console.log('Custom JavaScript loaded.');
      </script>
---

## Part 2: Coding in action

### Instructions

Please answer these questions in R using whatever syntax makes sense to you. Feel free to use the R help pages and/or notes from class.

**If you use Google or ChatGPT etc. to help you get to an answer, please note so on your answer**

Step 1. ARE YOU IN A BRANCH?????\
Step 2. Save this file as lastname_quiz.qmd in the same quiz folder.\
Step 3. Once you are done, click "render" above. This will create the .html file that I will evaluate.\
Step 4. Stage and commit your changes, push to your github.\
Step 5. Submit a pull request. If github warns you of conflicts please try and fix them before submitting your pull request (did you rename the file??? ).\
(Step 6. In case of utter disaster, email Sabrina your .html file.)

```{r, message = F}
# load packages and data 
library(ggplot2)
library(dslabs)
library(dplyr)
```

#### Load your data:

Load the *swallows.csv* into R, call it swallows. This data frame lists results of an experiment on tree swallows. Each row is a nest. Nests were put into one of two **treatments**: simulated predation (predator), or control (control). Along with treatment you have the following columns:\
**nest_fate** = whether the nest fledged any young or whether they all died\
**brood** = number of nestlings\
**n_fledged** = number of nestlings that fledged.

You also have a separate data frame called *brightness.csv* where each row lists the plumage brightness (a sexual signal) for the female at each nest. Load this one up too.

```{r}
# Type your answer in this box using R code 
swallows <- read.csv("swallows.csv", header = T, sep = ",")
brightness <- read.csv("brightness.csv", header = T, sep = ",")

head(swallows)
head(brightness)

# Google? y/n
# No
```

#### Wrangle your data

You realize that you want to add brightness information onto your main swallows data frame. Join these data frames using nest_id as your key variable. Describe how you would check to make sure this join worked as intended.undefined

```{r}
#Answer
# Joining the brightness df to the swallows df using left_join (swallows df in the left)

swallows2 <- left_join(swallows, brightness, by = join_by(nestbox))

# Checking if the brightness column was added
str(swallows2)

# Picking a random nest id to confirm that they have the correct brightness value in the new df
brightness %>% filter(nestbox == "2_95")
swallows2 %>% filter(nestbox == "2_95")

#Google? y/n
# The Help tab in R because I couldn't remember how to specify the column to join by
```

Inspect your combined data frame. Do some sort of quick sanity check on each column to catch any obvious entry errors. Correct as necessary

```{r}
# Answer
summary(swallows2)
filter(swallows2, is.na(brightness) == T)

table(swallows2$treatment)
table(swallows2$nestbox)
boxplot(swallows2$brightness)
boxplot(swallows2$brood)
boxplot(swallows2$n_fledged)


# There only seems to be one error, the NA value in brightness, but I didn't remove it because the rest of the data may be useful

# Google? y/n
# No
```

Create a new column called prop_fledged that contains the proportion of nestlings that fledged from each nest.

```{r}
#Answer

swallows2$prop_fledged <- swallows2$n_fledged/swallows2$brood

# Google? y/n
# No
```

#### Explore the data

What was the mean proportion of nestlings fledged in each treatment?

```{r}
# Answer 
mean(swallows2$prop_fledged, na.rm = T)

# Google? y/n
# No
```

Create a plot showing fledging success differences between treatments

```{r}
# Answer
ggplot(swallows2, aes(x = treatment, y = prop_fledged, fill = treatment)) +
  geom_boxplot() +
  labs(x = "Treatment", y = "Fledging success") +
  theme_bw()

# Success in the control is slightly higher compared to the predation treatment

# Google? y/n
#No
```

Is there any evidence that female brightness influenced fledging success? Create a plot to visually investigate this question.

```{r}
# Answer

ggplot(swallows2, aes(x = brightness, y = prop_fledged)) +
  geom_point() +
  labs(x = "Femle brightness", y = "Fledging success") +
  theme_bw()

# There doesn't seem to be any indication that female brightness is related to fledging success

# Google? y/n
# No
```
