---
title: "Quiz 1: Part 2"
format: html
editor: visual
embed-resources: true
css: |
      <style>
        body { font-family: Arial, sans-serif; }
      </style>
js: |
      <script>
        console.log('Custom JavaScript loaded.');
      </script>
---

## Part 2: Coding in action

### Instructions

Please answer these questions in R using whatever syntax makes sense to you. Feel free to use the R help pages and/or notes from class.

**If you use Google or ChatGPT etc. to help you get to an answer, please note so on your answer**

Step 1. ARE YOU IN A BRANCH?????\
Step 2. Save this file as lastname_quiz.qmd in the same quiz folder.\
Step 3. Once you are done, click "render" above. This will create the .html file that I will evaluate.\
Step 4. Stage and commit your changes, push to your github.\
Step 5. Submit a pull request. If github warns you of conflicts please try and fix them before submitting your pull request (did you rename the file??? ).\
(Step 6. In case of utter disaster, email Sabrina your .html file.)

```{r, message = F}
# load packages and data 
library(ggplot2)
library(dslabs)
library(dplyr)
```

#### Load your data:

Load the *swallows.csv* into R, call it swallows. This data frame lists results of an experiment on tree swallows. Each row is a nest. Nests were put into one of two **treatments**: simulated predation (predator), or control (control). Along with treatment you have the following columns:\
**nest_fate** = whether the nest fledged any young or whether they all died\
**brood** = number of nestlings\
**n_fledged** = number of nestlings that fledged.

You also have a separate data frame called *brightness.csv* where each row lists the plumage brightness (a sexual signal) for the female at each nest. Load this one up too.

```{r}
# Type your answer in this box using R code 
swallows <- read.csv('swallows.csv')
brightness <- read.csv('brightness.csv')

# Google? y/n
#n
```

#### Wrangle your data

You realize that you want to add brightness information onto your main swallows data frame. Join these data frames using nest_id as your key variable. Describe how you would check to make sure this join worked as intended.undefined

```{r}
#Answer
full_df <- swallows %>% 
  full_join(brightness, by=join_by(nestbox==nestbox))#, keep = T)

# Initially using the keep=True argument in full_join function allows for checking that joined dataframe properly matched each key

#Google? 
#n


```

Inspect your combined data frame. Do some sort of quick sanity check on each column to catch any obvious entry errors. Correct as necessary

```{r}
# Answer
full_df <- na.omit(full_df)
class(full_df$nestbox) # familiarize with datatypes across all columns

full_df$nestbox <- as.factor(full_df$nestbox)
full_df$treatment <- as.factor(full_df$treatment) # See that we are plotting later on, so just convert to factor now

length(unique(full_df$nestbox)) # should have 41 unique boxes and only two treatments

quantile(full_df$brood) # check that numbers for brood, n_fledged, brightness don't look unreasonable (no weird outliers that may be misinputs)

unique(full_df$brood >= full_df$n_fledged) # The number of nestlings that fledged should never exceed the initial brood number. If we get False here, something is wrong.

# Google y/n 
#n
```

Create a new column called prop_fledged that contains the proportion of nestlings that fledged from each nest.

```{r}
#Answer
full_df <- full_df %>% 
  mutate(prop_fledged = n_fledged/brood)

# There are NA values here due to divisions by zero in brood number. Will omit these later

# Google? y/n
#n
```

#### Explore the data

What was the mean proportion of nestlings fledged in each treatment?

```{r}
# Answer 

full_df %>% 
  group_by(treatment) %>% 
  summarise(mean(na.omit(prop_fledged))) # Just omiting NA's here 


# Google? y/n
#n
```

Create a plot showing fledging success differences between treatments

```{r}
# Answer
full_df %>% 
  ggplot(aes(x=treatment, y=prop_fledged)) +
  geom_boxplot() +
  labs(title = 'Fledging Success Across Treatments', y='Proportion of Nestlings Fledged Per Nest',
       x='')

# Google? y/n
#n


```

Is there any evidence that female brightness influenced fledging success? Create a plot to visually investigate this question.

```{r}
# Answer
full_df %>% 
  ggplot(aes(x=brightness, y=prop_fledged)) +
  geom_point() +
  labs(title = 'Effects of Female Plumage Brightness on Fledging Success', y='Proportion of Nestlings Fledged', x='Female Brightness')

# Google? y/n
#n

```

This data does not support the hypothesis that female plumage brightness has an effect on nestling fledging because there is no discernible relationship between these features in our scatter plot.
